---
title: "PRACTICE_baseball_data"
author: "Andrew Crena"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## R Markdown
##### I prefer to set wd's through the user interface by either using the upper left hand quadrant, or the bottom right.
```{r setup, include=TRUE, echo=FALSE}
knitr::opts_chunk$set(root.dir = "C:\\Users\\adcre\\OneDrive\\Documents\\Desktop_RStudio")
getwd()
```

## Baseball Data
##### Example of me importing a baseball data set, cleaning anything that needs it, and performing some exploratory analysis. Hopefully we will be able to use visualization, but my priority with this file is just to practice importing!

## Import from loacal files, still zipped
```{r from local, include=T, echo=T}
unzip_baseball <- unzip("C:\\Users\\adcre\\OneDrive\\Documents\\Desktop_RStudio\\baseball_data_zipped.zip")
baseball_data <- read.csv(unzip_baseball)

View(baseball_data)
```

## Import from URL
```{r import url, echo=T, warning=FALSE, message=FALSE, include=T}
library(utils)

# assign a variable to your url for future ref
url <- "https://www.kaggle.com/datasets/mathchi/hitters-baseball-data/download?datasetVersionNumber=1"
download.file(url, destfile = "dataset.zip")
unzpd_baseball <- unzip("dataset.zip")
baseball_data_url <- read.csv("C:\\Users\\adcre\\OneDrive\\Documents\\Desktop_RStudio\\Hitters.csv")

View(baseball_data_url)
```

## Clean/Subset
##### Let's see how we can subset this data. But first, take a look at your variables !
```{r clean and subset, include=T, echo=T}
baseball_data <- read.csv("Hitters.csv")
Column_indexes <- colnames(baseball_data)
Column_indexes

# It looks like we have to create a subject column, here is one way of doing this
baseball_data$subject_id <- row.names(baseball_data)
View(baseball_data)

# Let's subset players with over 200 at-bats
manyabs_data <- baseball_data[baseball_data$AtBat > 200, ]
View(manyabs_data)

# After looking through the data, there appears to be some missing values, especially in certain columns. Let's check and make a new variable that has a df only non-NA-containing cases
noNA_baseball <- subset(baseball_data, !is.na(baseball_data$Salary))
View(noNA_baseball)

# Let's combine these two conditions for our next subset, just to show how specialized you can make your calls for subsetting
new_baseball <- baseball_data[baseball_data$AtBat>250 & baseball_data$HmRun<5, ]
View(new_baseball)

# After observing this graph, I see there are columns for 'Hits' and 'AtBats', but no batting average column. This is a good opportunity to calculate batting average (Hits / AtBats), and create a new column that represents it on our data structure
baseball_data$batting_avg <- ((baseball_data$Hits)/(baseball_data$AtBat))
which(colnames(baseball_data) == "batting_avg")
bat_avg <- baseball_data[ ,22]

# Removing outliers
# grep() is being used here to remove an outlier, assuming we know the characters/numbers etc of the
# data. For the # sake of example, I will remove a case where a hitter had not enough at-bats (19) for 
# me to feel confident in their data. 
badid <- grep('65', baseball_data$subject_id)
baseball_data_sub <- baseball_data[-badid,]
View(baseball_data_sub)
```

##### Thankfully, baseball has become a very data-oriented game in the past decade. This allows us to have descriptive equations or statistics that can be easily visualized. Let's find a different baseball dat set online to practice on.
```{r more explore, include=T, echo=T}
# download new data set. There are two csv files I will be looking at. One will be hitting stats from the 
# postseason in the past half-century for all players that played in the playoffs, as well as another 
# data set that contains pitching stats. Let's se what we can learn!
hitting_data <- read.csv("BattingPost.csv")

# This is clearly too much, so let's just extract the last three years, 2012-2015 (info only kept until 
# 2015)
new_hitting <- hitting_data[hitting_data$yearID > 2012, ]
str(new_hitting)

# Now, let's use the info given in the columns to create a new column for 'slugging percentage' a 
# baseball stat that is meant to reflect how damging it is when a player does get a hit. For example, 
# Barry Bonds had a high Slugging percentage because he did damage much of the time when we successfully 
# got a hit. I'll give you a hint; Barry did not like hitting singles!
# 
# Per the MLB website, the formula for slugging percentage is: (1B + 2Bx2 + 3Bx3 + HRx4)/AB. 
new_hitting$slugging <- (((new_hitting$H) + ((new_hitting$X2B)*2) + ((new_hitting$X3B)*3) + ((new_hitting$HR)*4)) / new_hitting$AB)
View(new_hitting)

# After observing how many NAs were in my new slugging column, I will remove them so we just have cases where sluging is available.
noNA_slug <- subset(new_hitting, !is.na(new_hitting$slugging))
View(noNA_slug)
```

### DURING-Session REFLECTION
#####
### POST-Session REFLECTION
##### TBD..